The \project{unRAVE} catalog

### 2016-08-23 article/unrave.tex
Show git hash when compiling, or default to unknown if vc.tex doesn't exist

\newcommand{\githash}{UNKNOWN}
%\input{vc.tex}

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

@@

Here we cross-match our results against literature samples to verify that
our labels are accurate.  Because this work is an independent analysis, our
first point of reference is against the official \rave\ data releases.  These
comparisons are shown in Figures \ref{fig:rave-dr4-comparison} and 
\ref{fig:rave-dr5-comparison}, where we show our labels with respect to the
results in the fourth and fifth \rave\ data releases, respectively.  In order
to provide a fair comparison we only show stars that meet a number of quality
flags in both samples. 
\stub{What selections qualify as a fair comparison in both samples?}
There is very good agreement in $\teff$, with a bias
and RMS of just XX~K and XXX~K, respectively.  We find our $\logg$ values are
closest to the `calibrated' \rave\ surface gravities, which make use of 
asteroseismic information to correct (increase) the $\logg$ values along the
giant branch.  The calibrated \rave\ surface gravities also correct the 
location of the red clump (by $\sim0.5$~dex) to be in the same position in our
results.  However, while our $\logg$ values generally agree with the calibrated
\rave\ values along the giant branch, our $\logg$ values do differ slightly 
along the main-sequence.  This is principally because the main-sequence in 
this work tends to taper down towards higher $\logg$ values at cooler 
temperatures, whereas the comparison sample tend to have a slightly flatter
main-sequence.  This difference is not likely to have a very significant 
effect on the detailed abundance or spectrophotometric distance determinations
between these studies.

@@

After verifying that our stellar labels are comparable with other
high-resolution studies, here we verify that the abundance labels we find
are consistent with expectations from astrophysics.  One example is stars in
globular clusters, which show anti-correlations in light element abundances 
\citep[e.g.,][and references therein]{Da_Costa,Carretta_2009}.  In the \rave\
survey we found 70 stars with positions and radial velocities that are
consistent with being members of globular clusters.  Specifically we identified
49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
globular cluster NGC~3201, and 10 members of NGC~362.  
\stub{Were any of these stars unintentionally in the training set?}
All of these systems are well-studied and exhibit abundance anti-correlations 
in Na--O and Mg--Al \citep{people}.  Other correlations are also present 
(e.g., C--N), but those relationships are less relevant for this work because 
we do not measure C and N abundances here.

@@

% Some literature notes:
% See http://iopscience.iop.org/article/10.1088/0004-637X/731/1/64/pdf
% NGC 3201 has Na-O anti-correlation: Carretta et al. (2009)
% NGC 362 has "two discrete groups along the Na-O anti-correlation": https://arxiv.org/abs/1307.4085
% NGC 362 has Mg-Al relationship (low Mg scatter, large Al scatter) --> Carretta (2013)
% NGC 3201 has Mg-Al relationship (http://arxiv.org/pdf/1305.3645.pdf and references therein)

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

\begin{thebibliography}{dummy}

@@

>>>>

\IfFileExists{vc.tex}{\input{vc.tex}}{\newcommand{\githash}{UNKNOWN}}

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni. We note that the while Ti can be measured, its values in both 
APOGEE DR12 and DR13 are known to not follow the expected Galactic trends \citep{Holtzman_2015, Hawkins_2016}.
This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

@@

Here we cross-match our results against literature samples to verify that
our labels are accurate.  Because this work is an independent analysis, our
first point of reference is against the official \rave\ data releases.  These
comparisons are shown in Figures \ref{fig:rave-dr4-comparison} and 
\ref{fig:rave-dr5-comparison}, where we show our labels with respect to the
results in the fourth and fifth \rave\ data releases, respectively.  In order
to provide a fair comparison, we only show stars that meet a number of quality
flags in both samples. 
\stub{What selections qualify as a fair comparison in both samples?}
There is very good agreement in $\teff$, with a bias
and RMS of just XX~K and XXX~K, respectively.  We find our $\logg$ values are
closest to the `calibrated' \rave\ surface gravities, which make use of 
asteroseismic information to correct (increase) the $\logg$ values along the
giant branch.  The calibrated \rave\ surface gravities also correct the 
location of the red clump (by $\sim0.5$~dex) to be in the same position in our
results.  However, while our $\logg$ values generally agree with the calibrated
\rave\ values along the giant branch, our $\logg$ values do differ slightly 
along the main-sequence.  This is principally because the main-sequence in 
this work tends to taper down towards higher $\logg$ values at cooler 
temperatures, whereas the comparison sample tend to have a slightly flatter
main-sequence.  This difference is not likely to have a very significant 
effect on the detailed abundance or spectrophotometric distance determinations
between these studies.

@@

After verifying that our stellar labels are comparable with other
high-resolution studies, here we verify that the abundance labels we find
are consistent with expectations from astrophysics.  One example is stars in
globular clusters, which show anti-correlations in light element abundances 
\citep[e.g.,][and references therein]{Da_Costa,Carretta_2009}. In the \rave\
survey, \cite{Anguiano_2015} found 70 stars with positions and radial velocities that are
consistent with being members of globular clusters.  We have used that sample
Specifically, this GC sample includes 49 stars belonging to NGC~5139 ($\omega$~Centauri), 11 members of the retrograde
globular cluster NGC~3201, and 10 members of NGC~362.  
\stub{Were any of these stars unintentionally in the training set? \\ KH: Not using my training sets.}
All of these systems are well-studied and exhibit abundance anti-correlations 
in Na--O and Mg--Al \citep{people}.  Other correlations are also present 
(e.g., C--N), but those relationships are less relevant for this work because 
we do not measure C and N abundances here.

@@

% Some literature notes:
% See http://iopscience.iop.org/article/10.1088/0004-637X/731/1/64/pdf
% NGC 3201 has Na-O anti-correlation: Carretta et al. (2009)
% NGC 362 has "two discrete groups along the Na-O anti-correlation": https://arxiv.org/abs/1307.4085
% NGC 362 has Mg-Al relationship (low Mg scatter, large Al scatter) --> Carretta (2013)
% NGC 362 has a large metallicity spread in RAVE (Anguiano_2015) but not the case in high-res studies (e.g. https://arxiv.org/abs/1307.4085 has a metallicity spread of 0.05/0.08 dex) may indicate better cluster selection needed and can be done with cannon results.
% NGC 3201 has Mg-Al relationship (http://arxiv.org/pdf/1305.3645.pdf and references therein)

@@

\acknowledgements
A.~R.~C. warmly thanks Jonathan Bird (Vanderbilt) and Sven Buder (MPIA).
This research made use of: Astropy, a community-developed core Python package for
Astronomy \citep{astropy}, NASA's Astrophysics Data System Bibliographic Services;
and \project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for KH has been provided through the Simons Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

\begin{thebibliography}{dummy}
\bibitem[Anguiano et al.(2015)]{Anguiano_2015} Anguiano, B., Zucker, D.~B., Scholz, R.-D., et al.\ 2015, \mnras, 451, 1229

@@

\bibitem[Hawkins et al.(2016)]{Hawkins_2016} Hawkins, K., Masseron, T., Jofre, P., et al.\ 2016, arXiv:1604.08800

\bibitem[Holtzman et al.(2015)]{Holtzman_2015} Holtzman, J.~A., Shetrone, M., Johnson, J.~A., et al.\ 2015, \aj, 150, 148
